Rotation Period Evolution in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking

### 2019-07-01 Paper/Response1/response.txt

Dear Dr. Shafter,

Thank you for considering our manuscript. Below you will find our response to the referee's report where I list the issues raised in the report, denoted by REF with the referee's text in quotes for clarity, and how we addressed them in our revised manuscript, denoted by REPLY. Following the referee's comments and suggestions, we have substantially edited portions of the Results section to improve the concision and clarity of our work. These edits merely change the language and phrasing and do not alter the scientific content or our conclusions that were present in the original submitted manuscript. Note that in the revised manuscript, all text changes, except for deletions, are in bold text for emphasis. For each updated figure, we included "Updated Figure" in bold in the figure caption for clarity.

Additionally, we added citations for VPLanet, the software used to perform all of our simulations, to the "Software" section at the end of the manuscript. We did not included STELLAR here as it is a submodule of VPLanet and can only be used as a component of the VPLanet executable when performing our simulations.

Best,

David Fleming, Rory Barnes, Jim Davenport, and Rodrigo Luger

Response to Reviewer
--------------------

Dear Reviewer,

Thank you very much for your careful review of our manuscript.  Your in-depth comments and critiques pointed out weaknesses in our work that, once addressed, improved the quality of our work. Below, I list the issues raised in your report, denoted by REF with your text in quotes, and how we addressed them in our revised manuscript, denoted by REPLY. Note that in the revised manuscript, all text changes, except for deletions, are in bold text for emphasis. For each updated figure, we included "Updated Figure" in bold in the figure caption for clarity.

During the review process, we noticed that an erratum was recently released for the Matt et al. (2015) magnetic braking model, Matt et al., (2019). The erratum impacted the torque normalization in equation 3 in our paper, overestimating it by a factor of 3/2. This error impacted every simulation in our paper that used the Matt+2015 model. We implemented this fix in our code, re-ran all of the simulations, and found that this error had no qualitative impact on our results: we still find that binaries can rotate subsynchronously at short Porb and tidally-lock at long Porb, even out to 100 d. The error weakened the torque due to magnetic braking relative to tidal torques, making it slightly easier for tides to tidally-lock binaries, strengthening our results, in a sense. The number of binaries tidally-locking at long Porb did decrease somewhat, however, as strong spin-down torques due to the combination of magnetic braking and tides are required to spin down long Porb (Porb > 60 d) stars enough to get captured into the tidally-locked state. This decrease, however, was small and did not significantly impact our results. After re-running all our simulations, we re-generated every figure that displayed simulation results that used the Matt+2015 model and updated every quantity derived from our simulation results in the revised manuscript.

Referee Report + Responses
--------------------------

Major points

REF: "As mentioned above, the science presented in the paper is good enough to be published, the lack of concision and proof-reading makes difficult for the reader to estimate the paper at its true value. Many parts can be reduced to make the point of the paper clearer and the reading more convenient. I believe this will help a lot for the acceptance of this work in the community."

REPLY: We apologize for the deficient proofreading and lack of concision in the submitted manuscript. After making the edits to respond to your comments, we thoroughly proofread the manuscript and worked to make our language more direct and concise, making many edits throughout the paper, all in bold.

REF: "Even though the whole range of parameter is explored for equilibrium tides, nothing is said about the importance that dynamical tides could have on the dissipation. Yet, Ogilvie (2007,2013) showed they are a major element of tidal evolution of rotating stars and Mathis (2015) gave a formalism that could be easily implemented in this type of study. In particular, in the case of low-mass stars, it has been shown by eg Gallet & al. (2017) that the tidal dissipation factor (in the CPL formalism) will be strongly affected. I do not expect that the authors re-do all the models for this paper (though I would highly recommend that they do such calculations for a new paper), but the authors need to acknowledge the limitation of their model, at least in the opening section at the end, at best in the description of their model."

REPLY: Thank you for pointing this out.  Tidal dissipation due to the dynamical tide is an important physical consideration that can impact the rotation rates of low mass stars, e.g. in the star-planet tidal interactions examined in Bolmont et al. (2016). To address this point, we added the following text in bold in a new subsubsection, 2.2.5, in the description of our model:

An additional mechanism for tidal dissipation in low-mass stellar binaries not considered in this work is the dynamical tide. Dynamical tidal dissipation arises from the turbulent viscous damping of inertial waves that are excited in the stellar convective envelope by a tidal perturber, with Coriolis acceleration serving as the restoring force (Zahn 1975; Ogilvie+2007). Under the dynamical tide formalism, the stellar mass, evolving stellar structure, rotation rate, and tidal forcing frequency can all strongly impact the strength of tidal dissipation, which can span many orders of magnitude (Ogilvie+2007, Ogilvie 2013, Mathis 2015, Gallet+2017}. For example, adopting the tidal frequency-averaged model for tidal dissipation of Ogilvie (2013), both Mathis (2015) and Gallet+2017 show that dynamical tidal dissipation is enhanced on the pre-main sequence via the expansion of the stellar radiative core and rapid rotation, whereas the magnetic braking-driven spin-down on the main sequence decreases the tidal dissipation. Dissipation due to the dynamical tide could be important for some of the systems considered in this work since, for coplanar binary stars on circular orbits, inertial waves are excited in the stellar convective envelopes for P$_{orb} > $P$_{rot}/2$ and can drive significant spin and orbital evolution (e.g. Witte+2002, Ogilvie+2007, Bolmont+2016). Although semi-analytic models for dynamical tidal dissipation that account for the evolving stellar structure and rotation exist (e.g. Mathis 2015, Bolmont+2016, Gallet+2017), we do not consider them here as they are currently limited to circular orbits. We instead focus on exploring the impact of the equilibrium tide across a wide range of parameter space and leave an examination of how the combination of the dynamical and equilibrium tide impacts the rotation period evolution of low-mass binary stars for future work.

REF: "It can be interesting to compare the effects of different magnetic braking models, and I understand that it allows a connection to the computations done in Fleming et al. (2018) using the same torque. But the rotation evolution trends provided by Reiners & Mohanty (2012)'s paper have been discussed at length in the literature as not realistic. They did a great work and were the first to show a direct dependence to the Rossby number and a cluster comparison for different mass at the same time but their converged sequence for solar-like stars (1.0-0.4 M_\odot) is evolving in the wrong way with higher mass stars spinning slower once converged (see eg their fig 4 for Hyades/Praesepe). Moreover, this comparison is not mentioned in the abstract nor discussed in the conclusions. I would remove this part that doesn't seem to be needed for this paper and just add a sentence about it with something like "we tested another braking law and the conclusion's were no different"."

REPLY: Thank you for pointing this discussion out and explaining the short-comings of the model. As per your suggestion, we removed our examination and discussion of simulations using the Reiners & Mohanty (2012) model. Specifically, we removed Sections 4.1.3 and 4.4.2 and Figures 5 and 14 in the original submitted manuscript.

REF: "Abstract : CPL is not define"

REPLY: The CPL model is now properly defined in the abstract.

1.Introduction

REF: "-Reiners & Mohanty (2012) didn't explore the Kepler field but did comparison with Praesepe"

REPLY: We corrected this errant citation.

REF: "Gyrochronlogy => Gyrochronology (A lot of typos are present all over the text, a bit of proof reading would be nice)"

REPLY: We corrected these typos in the text and apologize for the lack of sufficient proofreading in the initial submission.

REF: "Abt & Booyarak (2004) found that long periods binary's Prot can deviate from that of single stars was applicable for A and B type stars. Is that expected to be valid for lower masses ?"

REPLY: Thank you for pointing this out.  After reviewing the literature, we believe that that connection is a bit too speculative as tidal processes in early and late type stars are likely not similar due to the different stellar structures, e.g. radiative vs convective envelope. Mazeh (2008) explained how Zahn (1975) showed that the tidal orbital circularization timescales differ between stars with radiative and convective envelopes, so we would anticipate that such a difference exists between the synchronization timescales as well. We have therefore removed the reference to Abt & Booyarak (2004) in both the introduction and in Section 4.3 as it is not applicable as we had originally believed.

REF: "Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyro chronology could be incorrect owing to the influence of tidal torques. No previous study has quantified this effect". Indeed, it would be a really nice output of that paper or a following one."

REF: "4.4 A lot can be cleaned here. It would have been more relevant to mimic a realistic population of binaries in the field to estimate the impact of binarity on gyrochronology."

REPLY: We grouped these comments as they touched on a similar point. Thank you for pointing out this deficiency in our results: we did not properly quantify the impact of tidally-interacting binaries on age estimates using gyrochronology. To address this deficiency, we created a new figure, Figure 13 in the revised manuscript, that quantifies the systematic error introduced by applying gyrochronology methods to tidally-interacting binaries. To quantify this error, we computed the the difference between the mean ages of single stars and tidally-interacting binaries in mass - Prot bins for both the CPL and CTL model simulation suites. We defined "tidally-interacting binaries" as binaries that were tidally-locked or had Prots within 10% of Peq, i.e. we combined the "Interacting" and "Locked" populations examined in Figures 9 and 10 in the revised manuscript. We found that gyrochronology methods can underpredict ages by up to 300% at short Prot and overpredict ages by up to 25% at long Prot. Only tidally-interacting binaries populated the Prot < 10 d and > 50 d bins for solar mass stars, so we could not compare to the ages of single stars in these cases. We expanded on these points in more detail and significantly edited Section 4.4, "Deviations From Single Star Prot Evolution: Implications for Gyrochronology", in the revised manuscript to accommodate these changes and clean up the section to improve the concision of our writing.

2. Methods

REF: "STELLAR is not a stellar evolution model. Be careful with the vocabulary used here, at best it can be considered as an interpolation tool in Baraffe+98's tracks. (Same comment at the beginning of the last paragraph of 2.1)"

REPLY: Thank you for pointing out this error in our description of the model. We have corrected this error in both locations in the text.

REF: "I assume it is only a typo in the text but an exponent is missing in the descriptions of Matt+15 magnetic braking in eq. 2 for the convective turnover timescale."

REPLY: We corrected this typo and confirmed that our code used the correct exponent in all simulations.

REF: "2.2.3 (end) What prevents your model from capturing other spin-orbit resonances ?"

REPLY: The CPL model, owing to its inclusion of only 4 discrete tidal lags, only allows a body to enter into 3:2 and 1:1 spin - orbit commensurabilities. The CTL model does not resolve capture into spin-orbit resonances besides 1:1 since it predicts a continuous equilibrium period as a function of the orbital period, eccentricity, and obliquity, for tidally-locked bodies. See Sections 2.1 and 2.2 in Barnes (2017) for an expanded discussion. Models that do resolve captures into spin-orbit resonances, e.g. Rodriguez et al. (2012), instead directly integrate the equations of motion and torques for tidally-interacting bodies, whereas our model is purely secular.

3. Simulations

REF: "I know it is not expected to evolve very much in that direction but what happened in the eccentricity increases and overcomes 0.3 in the CPL case ?"

REPLY: In the CPL model, the eccentricity can grow for certain orbital and rotational states (see Equation 8 in Barnes 2017), such that the eccentricity could increase to beyond 0.3. In such a case, the CPL model can still be applied, however, it is not strictly valid as it is in effect derived by truncating the equations for tidal evolution to second order in eccentricity, whereas the CTL model is 8th order in eccentricity. Leconte et al. (2010) argued that this truncation renders the predictions of the CPL model incorrect in this eccentricity regime. Specifically, Leconte at al. (2010) found that for e > 0.32, the higher-order eccentricity terms dominate the constant and second order terms, meaning the the CPL model does not accurately capture the physics of the tidal interaction for moderate to large eccentricities. In this case, Leconte et al. (2010) showed that this can lead to semi-major axis growth under the 2nd order CPL model instead of the decrease predicted by the CTL model. Moreover, Leconte et al. (2010) found that under the CPL model for pseudo-synchronously (supersynchronously) binaries with eccentricities > 0.326, the eccentricities can grow, whereas the CTL model predicts that the eccentricity should damp. Leconte et al. (2010) argue that the latter is the correct behavior. We added a brief reference to this discussion in Leconte et al. (2010) in the text to clarify this point.

REF: "Respectfully" -> Respectively (and elsewhere in the text)"

REPLY: We have corrected all of these errors in the text.

4. Results

REF: "4.1.1 Low-mass stars can be in the saturated regime for very slow rotation rates + In your range of parameter, a binary 1Msun-1Msun is an extreme case with maximal tidal effect. So it doesn't seem very appropriate for a representative example as fig. 3."

REPLY: Those are good points, however, we chose to use a 1Msun - 1 Msun binary as a pedagogical case because even with the extreme tidal torques in such a binary, tides and magnetic braking can still balance to produce subsynchronous rotation, similar to how we consider short Porb systems where tides are strong and in addition to long Porb cases up to 100 d. Furthermore, given our assumptions and under the Matt+2015 magnetic braking model, both stars would be in the unsaturated regime, justifying our assumptions. We clarified these assumptions in Section 4.1.1 in the revised manuscript and by adding the following text:

Although solar mass stars are the most massive stars we consider in this work, and hence will have the strongest tidal torque for a given tidal dissipation parameter and P$_{orb}$, they can still exhibit subsynchronous rotation and serve as a useful pedagogical case to examine here and in simulations below.

Additionally, we took this opportunity to update Fig. 3 by making the range of k2 * tau values more symmetric about our fiducial values and added contours for clarity.

REF: "4.1 is very interesting and the core of the paper. It can however be improved with more concision."

REPLY: We edited and trimmed down Section 4.1 to improve both the concision and focus. We bolded nearly all the text in Section 4.1 as most portions were edited to improve concision without any substantial nor qualitative change to the scientific content of the text.

REF: "4.1.4 would be more appropriate as part of section 4.2 in my opinion"

REPLY: Upon further reflection, we agree that 4.1.4 would be better suited as part of Section 4.2. It is now Section 4.2.1 in the revised manuscript and we have slightly edited the text near the end of Section 4.2 and beginning of 4.2.1 to accommodate its inclusion.

REF: "4.2 Fig 8 and 9 needs to be adjusted and reshaped, legend, white is not very visible on yellow, etc... They are both really interesting and I spent a lot of time understanding each part of the diagram but I was disappointed that no discussions were made about the narrowness and the wideness of some constraining intervals, particular around for the CPL model. Thus I advise either to remove the interval in the figure, that would allow it to be smaller and clearer and there is already a lot to say about the figure without them. Either to add some comment in the text in a short and concise way."

REPLY: We removed the intervals from the figure since displaying the median values still capture the physics of the various spin-orbital states examined in said figure. We significantly increased the font size to improve the figure's legibility.

REF: "4.3 It would be interesting given the type of diagram of figure 10 and 11, to compare to the Kepler rotation period field by McQuillan+2014 for example and to look at what could be some missed binaries."

REPLY: We are glad you think this as we believe it would be an interesting study as well! We presently are pursuing this very point in a separate project using the same model presented in the submitted manuscript. We presented this work-in-progress at the 2019 meeting of the American Astronomical Society (Here's an ADS link to the abstract for reference: http://adsabs.harvard.edu/abs/2019AAS...23334101F) and leave this for a future publication.

REF: "Also, I don't think anyone ever gave an age to a star with gyrochronology using Reiners&Mohanty (2012). I wouldn't recommend the use of the two different saturations value in a single population model."

REPLY: We have removed the discussion and analysis involving the Reiners & Mohanty (2012) model.

REF: "4.4.1 has actually been already discussed in 4.3."

REPLY: We have removed this subsubsection as it mainly repeated content from 4.3.

REF: "4.5 This section has to be present but should be twice smaller in my opinion and merged with 4.6 and 4.7, the message is very diluted here. Also, I'm surprised by the conclusion of the authors, while they claim that it is hard to make a clear statement, it seems very clear to me that the CTL model is doing a better job a reproducing the field by Lurie+2017. Particularly because the CPL models seems enable to produce super-synchronous rotators. But this is only an eye judgement. I'm also surprise to see that nothing is said about the highly super synchronous rotators with high eccentricity in Lurie+2017 sample. Are they still in the modelled range of parameters ? Should the authors expect their models to reproduce these binary ?"

REPLY: As addressed in your point below, we removed section 4.7 and compressed its point to a few sentences added to the Discussion. We wish to keep 4.5 and 4.6 separate as 4.5 deals with examining how our model predictions compare to the Lurie+2017 data whereas 4.6 offers general observational tests to distinguish between the CPL and CTL model.

We agree that qualitatively, the CTL model does a better job reproducing the Lurie+2017 data than the CPL model as the former populates regions of parameter space that the CPL model fails to do, e.g. the supersynchronous population noted by the reviewer. We choose to be cautious with our wording and stick with a more qualitative interpretation as our goal in this work was not to perform inference, but instead to examine the model predictions across a wide range of parameter space. To this end, we chose our tidal dissipation parameter and mass prior distributions to be intentionally broad such that they span a plausible region of parameter space. For inference and proper statistical model selection, we would have instead used a more proper mass prior, perhaps sampling the primary star masses, with replacement, from the masses in the Kepler Input Catalogue, or by using a reasonable initial mass function, but still accounting for Kepler's bias towards predominantly detecting FGK stars. Furthermore, since the underlying distribution of tidal dissipation parameters as a function of stellar mass and age is not well known, we did not believe that any inference between the CPL and CTL models we made would be robust. When performing inference, we could instead test the impact of different distributions of tidal dissipation parameters, or adopt a model for tidal dissipation as a function of stellar mass and age, like the Mathis (2015) model. Moreover, even though the Lurie+2017 data is currently the best available for such studies, its lack of uncertainties on Prot and eccentricties, and the approximate means in which the eccentricities were derived, prevent a robust comparison. We have significantly edited section 4.5, with most of the text reflecting the changes in bold in the revised manuscript, to compress the section, clarify the message, and address the differences between the CTL and CPL model predictions when compared to the Lurie+2017 data. We additionally worked to shorten the section and note that these edits did not significantly alter the scientific content nor conclusions.

As for the highly supersynchronous rotators, our CTL model could reproduce them if they are tidally-locked binaries with eccentricities > 0.3, beyond the range we considered. In this case for large eccentricities, the CTL model equilibrium period, our equation 22, predicts that binaries on more eccentric orbits rotate more supersynchronously, e.g. Fig. 1 in Barnes (2017), so if we considered more eccentric binaries, our model could reproduce that observation. We have updated the colorbar for Fig. 15 to note that the highly supersynchronous rotators in the Lurie+2017 data have eccentricities > 0.3 and added the following text discussing why our model does not reproduces these systems in the revised manuscripts:

### 2019-07-01 Paper/sync.tex
Updates from Overleaf

We examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries up to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters using two common equilibrium tidal models. We find that many binaries with P$_{orb} \lsim 20$ d tidally-lock, and most with $P_{orb} \lsim 4$ d tidally-lock into synchronous rotation on circularized orbits. At short P$_{orb}$, tidal torques produce a population of fast rotators that single-star only models of magnetic braking fail to produce.  In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, qualitatively reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}. Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the \xxx{Constant Phase Lag tidal} model predicts that binaries can tidally-lock out to P$_{orb} \approx 100$ d. Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e. gyrochronology \xxx{can underpredict ages by up to $300\%$} unless one accounts for binarity. We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.

 %Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries. Here we examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries out to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters. In many cases, we find that magnetic braking can overpower tides, producing a long-lasting population of subsynchronous rotators, even in short P$_{orb}$ binaries, reproducing the observed \kepler eclipsing binaries (EBs) population. Furthermore, at short P$_{orb}$, tidal locking naturally producing a population of fast rotators that cannot be reproduced by single-star-only models of magnetic braking.  Many binaries can tidally-lock even out to P$_{orb} \approx 100$ d, fixing P$_{rot}$ to P$_{orb}$, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e. gyrochronology methods must be applied carefully. We compare the predictions of two common equilibrium tidal models and show that the ``Constant Time Lag" equilibrium tidal model provides a better match to observations. We recommend observational tests to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars. Finally, by comparing observations of \kepler EBs with our simulations, we constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.

@@

The long-term angular momentum evolution of low-mass (M$\lsim 1$ M$_{\odot}$) stars is controlled by magnetic braking, the torque exerted on stars due to the coupling of stellar winds to the surface magnetic field \citep{Mestel1968}. Early in stellar lifetimes, stars spin-up as they contract along the pre-main sequence.  Once stars reach the main sequence, stellar radii remain mostly constant while magnetic braking removes angular momentum from the stars, gradually spinning them down over time \citep{Skumanich1972}. Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \textbf{\citep[e.g. Praesepe, ][]{Reiners2012,Matt2015,Douglas2017}} and field stars \citep[e.g. the \kepler field, ][]{Matt2015,vanSaders2018}. Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as \xxx{gyrochronology} \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.

@@

Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially for stellar spins. Tidal torques drive binaries towards the tidally-locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.  Tidal-locking occurs much earlier than orbital circularization with the tidal-locking timescale estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is typically much less angular momentum in stellar spins than the binary orbit. As a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.

In low-mass binaries, both magnetic braking and tidal torques compete to shape the stellar P$_{rot}$ evolution. When tides dominate, in particular at close orbital separations, tides can fix P$_{rot} =$ P$_{orb}$, or more generally P$_{rot} =$ P$_{eq}$ for eccentric orbits. In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally-locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2018}. Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally-locked state into subsynchronous rotation \citep[P$_{rot} >$ P$_{eq}$, ][]{Habets1989,Zahn1994,Keppens1997}. This behavior seems to be bourne out in nature, as \citet{Lurie2017} discovered a substantial population of subsynchronous short-period binaries in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\approx} 0.9$, in defiance of the expectation of tidal locking at such short orbital separations. \xxx{The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.}

% cut
% The influence of tides is not necessarily limited to short P$_{orb}$ systems, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered. 
%The large sample of $816$ P$_{rot}$ measurements derived using star spot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally-locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d. Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking. These findings suggest that synchronization is not a given for tidally-interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$. Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.

Understanding the interaction between tidal torques and magnetic braking is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g. Praesepe, ][]{Agueros2011,Douglas2017} and field stars \citep[e.g. \kepler, ][]{Reinhold2013,McQuillan2014} are likely contaminated by unresolved binaries given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys. In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.  \xxx{Tidally-interacting binaries impart a} contaminating signal that is not currently accounted for by models. Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology could be incorrect owing to the influence of tidal torques. No previous study has quantified this effect.

@@

We improve upon the \xxx{interpolation of the \citet{Baraffe2015} stellar evolution models employed by \citet{Fleming2018}, \stellar, by additionally performing a bicubic interpolation of the stellar radius of gyration, $r_g$, over mass and time of the \citet{Baraffe2015} models. This updated version of \stellar now tracks the full moment of inertia evolution of low-mass stars according to the \citet{Baraffe2015} stellar evolution models, a critical requirement for modeling the angular momentum evolution of low-mass stars.}

>>>>

We examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries up to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters using two common equilibrium tidal models. We find that many binaries with P$_{orb} \lsim 20$ d tidally lock, and most with $P_{orb} \lsim 4$ d tidally lock into synchronous rotation on circularized orbits. At short P$_{orb}$, tidal torques produce a population of fast rotators that single-star only models of magnetic braking fail to produce.  In many cases, we show that the competition between magnetic braking and tides produces a population of subsynchronous rotators that persists for Gyrs, even in short P$_{orb}$ binaries, qualitatively reproducing the subsynchronous eclipsing binaries (EBs) discovered in the \kepler field by \citet{Lurie2017}. Both equilibrium tidal models predict that binaries can tidally-interact out to P$_{orb} \approx 80$ d, while the Constant Phase Lag tidal model predicts that binaries can tidally lock out to P$_{orb} \approx 100$ d. Tidal torques often force the P$_{rot}$ evolution of stellar binaries to depart from the long-term magnetic braking-driven spin down experienced by single stars, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e. gyrochronology can underpredict ages by up to $300\%$ unless one accounts for binarity. We suggest that accurate determinations of orbital eccentricties and P$_{rot}$ can be used to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars.

 %Tidal torques dominate the angular momentum evolution of short-period, low-mass stellar binaries. Here we examine how tides, stellar evolution, and magnetic braking shape the rotation period (P$_{rot}$) evolution of low-mass stellar binaries out to orbital periods (P$_{orb}$) of 100 d across a wide range tidal dissipation parameters. In many cases, we find that magnetic braking can overpower tides, producing a long-lasting population of subsynchronous rotators, even in short P$_{orb}$ binaries, reproducing the observed \kepler eclipsing binaries (EBs) population. Furthermore, at short P$_{orb}$, tidal locking naturally producing a population of fast rotators that cannot be reproduced by single-star-only models of magnetic braking.  Many binaries can tidally lock even out to P$_{orb} \approx 100$ d, fixing P$_{rot}$ to P$_{orb}$, revealing that P$_{rot}$ is not be a valid proxy for age in all cases, i.e. gyrochronology methods must be applied carefully. We compare the predictions of two common equilibrium tidal models and show that the ``Constant Time Lag" equilibrium tidal model provides a better match to observations. We recommend observational tests to discriminate between which equilibrium tidal models best describes tidal interactions in low-mass binary stars. Finally, by comparing observations of \kepler EBs with our simulations, we constrain low-mass stellar tidal time lags to $\tau = 0.19^{+0.43}_{-0.16}$ s.

@@

The long-term angular momentum evolution of low-mass (M$\lsim 1$ M$_{\odot}$) stars is controlled by magnetic braking, the torque exerted on stars due to the coupling of stellar winds to the surface magnetic field \citep{Mestel1968}. Early in stellar lifetimes, stars spin-up as they contract along the pre-main sequence.  Once stars reach the main sequence, stellar radii remain mostly constant while magnetic braking removes angular momentum from the stars, gradually spinning them down over time \citep{Skumanich1972}. Although the precise details of how magnetic braking operates are not fully known, models of magnetic braking have been used to successfully model the bulk trends of P$_{rot}$ distributions in clusters \citep[e.g. Praesepe, ][]{Reiners2012,Matt2015,Douglas2017} and field stars \citep[e.g. the \kepler field, ][]{Matt2015,vanSaders2018}. Furthermore, the magnetic braking-driven long-term spin-down of stars has been used to estimate stellar ages, a method known as gyrochronology \citep{Skumanich1972,Barnes2003,Barnes2007,Mamajek2008,Barnes2010}, with older stars assumed to have lost more angular momentum due to magnetic braking and therefore rotate more slowly.

@@

Tides impart a significant signature in the long-term angular momentum evolution for binary stars, especially for stellar spins. Tidal torques drive binaries towards the tidally locked state in which the stellar P$_{rot}$ is equal to the equilibrium rotation period (P$_{eq}$) predicted by tidal models, with a familiar example of this effect being spin-orbit synchronization where P$_{rot} = $ P$_{eq} = $ P$_{orb}$.  Tidal-locking occurs much earlier than orbital circularization with the tidal-locking timescale estimated to be $2-3$ orders of magnitude less than the circularization timescale \citep{Zahn1989,Witte2002,Mazeh2008} as there is typically much less angular momentum in stellar spins than the binary orbit. As a result, tidal-locking is expected for binaries with P$_{orb} \lsim$ 20 d \citep[e.g.][]{Levato1974,Meibom2006,Mazeh2008,Zahn2008,Meibom2015}.

In low-mass binaries, both magnetic braking and tidal torques compete to shape the stellar P$_{rot}$ evolution. When tides dominate, in particular at close orbital separations, tides can fix P$_{rot} =$ P$_{orb}$, or more generally P$_{rot} =$ P$_{eq}$ for eccentric orbits. In such situations, magnetic braking still operates, removing angular momentum from each star, forcing tides to compensate for each star's loss of angular momentum by spinning up the stars to maintain the tidally locked equilibrium, removing angular momentum from the orbit, hardening the binary \citep[][]{Verbunt1981,Repetto2014,Fleming2018}. Tides do not win out over magnetic braking in general, however, as magnetic braking can spin-down the stars past the tidally locked state into subsynchronous rotation \citep[P$_{rot} >$ P$_{eq}$, ][]{Habets1989,Zahn1994,Keppens1997}. This behavior seems to be bourne out in nature, as \citet{Lurie2017} discovered a substantial population of subsynchronous short-period binaries in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\approx} 0.9$, in defiance of the expectation of tidal locking at such short orbital separations. The competition between magnetic braking and tidal torques can lead to complex angular momentum evolution in low-mass stellar binaries, and no previous work has conducted a systematic study to examine how this evolution proceeds across a wide range of tidal dissipation parameters and P$_{orb}$.

% cut
% The influence of tides is not necessarily limited to short P$_{orb}$ systems, as \citet{Abt2004} found that the P$_{rot}$ evolution of stellar binaries out to P$_{orb} \approx 500$ d can deviate from that of single stars, meriting the examination of the impact of tides out to longer P$_{orb}$ than is traditionally considered. 
%The large sample of $816$ P$_{rot}$ measurements derived using star spot modulations on low-mass \kepler eclipsing binary stars by \citet{Lurie2017} confirm that most short-period binaries are tidally locked and also show that the tidal torques can influence stellar spins in binaries with P$_{orb}$ up to 45 d. Futhermore, \citet{Lurie2017} shows that a substantial a population of subsynchronous short-period binaries exist in the \kepler field, clustered near P$_{orb}/$P$_{rot}{\sim} 0.9$, in defiance of the expectation of tidal locking. These findings suggest that synchronization is not a given for tidally interacting, short P$_{orb}$ stellar binaries and that impact of tidal effects should be reconsidered for longer P$_{orb}$. Both the extended \kepler mission \citep[K2,][]{Howell2014} and the Transiting Exoplanet Survey Satellite \citep[TESS, ][]{Ricker2014,Sullivan2015} will detect additional low-mass eclipsing binaries, permitting additional tests for models of stellar angular momentum evolution subject to tidal torques.

Understanding the interaction between tidal torques and magnetic braking is of paramount importance as P$_{rot}$ distributions measured in clusters \citep[e.g. Praesepe, ][]{Agueros2011,Douglas2017} and field stars \citep[e.g. \kepler, ][]{Reinhold2013,McQuillan2014} are likely contaminated by unresolved binaries given that roughly half of Sun-like stars are in stellar binaries \citep{Raghavan2010,Duchene2013}, and that binaries are difficult to resolve in photometric surveys. In the \kepler field, for example, \citet{Simonian2018} recently found that most rapid rotators (P$_{rot} \lsim 7.5$ d) are likely non-eclipsing, tidally-synchronized short-period photometric binaries, indicating that tidal torques in binaries can significantly impact observed P$_{rot}$ distributions.  Tidally-interacting binaries impart a contaminating signal that is not currently accounted for by models. Moreover, any ages inferred from rotation periods of stars in unresolved binaries using gyrochronology could be incorrect owing to the influence of tidal torques. No previous study has quantified this effect.

@@

We improve upon the interpolation of the \citet{Baraffe2015} stellar evolution models employed by \citet{Fleming2018}, \stellar, by additionally performing a bicubic interpolation of the stellar radius of gyration, $r_g$, over mass and time of the \citet{Baraffe2015} models. This updated version of \stellar now tracks the full moment of inertia evolution of low-mass stars according to the \citet{Baraffe2015} stellar evolution models, a critical requirement for modeling the angular momentum evolution of low-mass stars.
